%!TEX root = lucene4IR2016workshop_report.tex
\section{Keynotes and Invited Talks}
%<<<<<<< HEAD
%During the course of the workshops a series of talks were given on how Lucene is being used in industry, teaching and for evaluation along with more technical talks on the inner workings of Lucene's scoring algorithm as well as how learning to rank is being included into Solr\footnote{\scriptsize{Slides are available from \url{www.github.com/leifos/lucene4ir}}}. 
%
%\subsection*{Introduction Talk: Why are we here?}
%{\bf Leif Azzopardi, University of Strathclyde}:
%Leif explained how after attending the lively Reproducibility workshop~\cite{arguello2016repro} at ACM SIGIR 2015, he wondered where the Lucene team was, and if Lucene and the community is so big, why they don't come to IR conferences - he posited that perhaps we haven't been inclusive and welcoming as we could to such a large community of search practitioners. He further asserted that this has lead to a failure in transferring our knowledge into one of the largest open source toolkits available. He argued that if we as academics want to increase our impact then we need to improve how we transfer our knowledge to industry. One way is working with large search engines, but what about other industries and organisations that need search and use toolkits like Lucene? He argued that we need to start speaking the same language i.e. work with Lucene et al. and look for opportunities on how we can contribute and develop resources for training and teaching IR and how to undertake evaluations and data science using widely used, supported and commonly accepted Open Source toolkits. He described how this workshop was a good starting point and opportunity to explore how academia and industry can better work together, where we can identify common goals, needs and resources that are needed to foster this relationship. 
%=======
During the course of the workshops a series of talks on how Lucene is being used in Industry, Teaching and for Evaluation along with more technical talks on the inner workings of how Lucene's scoring algorithm works and how learning to rank is being included into Solr, were presented\footnote{\scriptsize{Slides are available from \url{www.github.com/leifos/lucene4ir}}}. A summary of each talk is below.

\subsection*{Introduction Talk: Why are we here?}
{\bf Leif Azzopardi, University of Strathclyde}:
Leif explained how after attending the lively Reproducibility workshop~\cite{arguello2016repro} at ACM SIGIR 2015, he wondered where the Lucene team was, and why, if Lucene and the community is so big, why they don't come to IR conferences - he posited that perhaps we haven't been very inclusive or welcoming to such a large community of search practitioners. He further asserted that this has reduced our capacity to transfer our knowledge and experience into one of the largest Open Source toolkits available. He argued that if we as academics want to increase our impact then we need to improve how we transfer our knowledge to industry. One way is working with large search engines, but what about other industries and organisations that need search and use toolkits like Lucene? He argued that we need to start speaking the same language i.e. work with Lucene et al and look for opportunities on how we can contribute and develop resources for training and teaching IR and how to undertake evaluations and data science using widely used, supported and commonly accepted Open Source toolkits. He described how this workshop was a good starting point and opportunity to explore how academia and industry can better work together, where we can identify common goals, needs and resources that are needed to foster this relationship. 
%>>>>>>> f0e54b9adc801bc5760e93a856bf677df92c95d5

\input{charlie}

\input{juanma}

\input{rup}

\input{jake}

\input{diego}
